---
title: "Extending GenomicDataStream"
subtitle: "File and data types"
author: "Developed by [Gabriel Hoffman](http://gabrielhoffman.github.io/)"
date: "Run on `r format(Sys.time())`"
output:
  rmarkdown::html_document:
    highlight: pygments
    toc: true
    toc_depth: 3
    fig_width: 5
#bibliography: library.bib
vignette: >
  %\VignetteIndexEntry{Extending GenomicDataStream}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\usepackage[utf8]{inputenc}
---

<!---


--->

```{r setup, echo=FALSE, results="hide"}
knitr::opts_chunk$set(
  tidy = FALSE,
  cache = TRUE,
  dev = "png",
  package.startup.message = FALSE,
  message = FALSE,
  error = FALSE,
  warning = FALSE
)
options(width = 100)
```	

<style>
body {
text-align: justify}
</style>

# Additional file types
Consider extending the package to support to support a new file of type `X`.  First, create a new class `Xstream` that inherits from `GenomicDataStream` defined [here](../doxygen/html/classgds_1_1_genomic_data_stream.html).  `Xstream` must define a constructor taking `Param` defined [here](../doxygen/html/structgds_1_1_param.html), and implement interfaces for `getNextChunk()` for each data type.

A [`DataChunk`](../doxygen/html/classgds_1_1_data_chunk.html) stores information about each feature and sample in a class inheriting from [`DataInfo`](../doxygen/html/classgds_1_1_data_info.html).  Streams for genotype data store feature information in the [`VariantInfo`](../doxygen/html/classgds_1_1_variant_info.html) class, while `DelayedStream` uses [`MatrixInfo`](../doxygen/html/classgds_1_1_matrix_info.html).  These classes or a custom class can be used to store information for a new file type. 

While nothing else about the underlying data form at is assumed, `GenomicDataStream` designed to chunks of _features_ rather than chunks of _samples_. Features are stored as columns in the matrix returned by R/C++, <ins>independent of the underlying data storage format</ins>.    


# Additional matrix types
First, take a look a currently supported [data types](./GenomicDataStream.html#data-types), and note that each of these types wraps an array of `double` storing data in column-major order.  In fact, the constructors to the dense matrix types for Eigen and Armadillo return objects that point to the original `double` array, without allocating new memory.  So any new data type should have a constructor that takes a `double` array.

The simplest new type to implement would be a `float` version of an existing type.  This could reduce memory usage in downstream analyses and increase speed at the cost of numerical precision.   






<!-- 
# Session info
<details>
```{r sessionInfo, echo=FALSE}
sessionInfo()
```
<\details> -->


<!-- # References -->

