---
title: "Issues on Mac OS X"
subtitle: "Installing TBB and RcppParallel"
author: "Developed by [Gabriel Hoffman](http://gabrielhoffman.github.io/)"
date: "Run on `r format(Sys.time())`"
output:
  rmarkdown::html_document:
    highlight: pygments
    toc: true
    toc_depth: 3
    fig_width: 5
vignette: >
  %\VignetteIndexEntry{Addressing issues on Mac OS X}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\usepackage[utf8]{inputenc}
---

<!---


--->

```{r setup, echo=FALSE, results="hide"}
knitr::opts_chunk$set(
  tidy = FALSE,
  cache = TRUE,
  dev = "png",
  package.startup.message = FALSE,
  message = FALSE,
  error = FALSE,
  warning = FALSE
)
options(width = 100)
``` 

<style>
body {
text-align: justify}
</style>

`GenomicDataStream` runs analyses across multiple cores in parallel using Threading Building Blocks in [oneTBB](https://uxlfoundation.github.io/oneTBB/).  This is available in R using [RcppParallel](https://cran.r-project.org/package=RcppParallel).  This package usually installs without an issue, but if it fails you can use the package after modifying the paths:


#### Install TBB using [brew](https://brew.sh)
```sh
brew install tbb
```

#### Install RcppParallel with TBB path
```r
DIR = system("echo $(brew --prefix tbb)", intern=TRUE)
TBB_INC = paste(DIR, "include/", sep='/')
TBB_LIB = paste(DIR, "lib/", sep='/')

.Internal(Sys.setenv("TBB", DIR))
.Internal(Sys.setenv("TBB_INC", TBB_INC))
.Internal(Sys.setenv("TBB_LIB", TBB_LIB))

install.packages("RcppParallel")
```



# Session info
<details>
```{r sessionInfo, echo=FALSE}
sessionInfo()
```
<\details>


<!-- # References -->

