---
title: "Demonstrate performance of [GenomicDataStream](https://gabrielhoffman.github.io/GenomicDataStream/)"
subtitle: 'Genetic variants'
author: "Developed by [Gabriel Hoffman](http://gabrielhoffman.github.io/)"
date: "Run on `r Sys.time()`"
output: 
  html_document:
    toc: true
    smart: true
    self_contained: true
---

<!---


# rm -rf demo_cache/

 rmarkdown::render("demo.Rmd");


--->


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  dev = c("png", "pdf"),
  package.startup.message = FALSE,
  cache = TRUE,
  cache.lazy = FALSE)
```

```{r getData, eval=FALSE}
wget https://ftp.1000genomes.ebi.ac.uk/vol1/ftp/data_collections/1000G_2504_high_coverage/working/20220422_3202_phased_SNV_INDEL_SV/1kGP_high_coverage_Illumina.chr21.filtered.SNV_INDEL_SV_phased_panel.vcf.gz

CHR=21
OUT=1kg_chr${CHR}

# VCF
bcftools view -i "(AF > 0.001) & (AF < 0.999) & (TYPE='snp')" 1kGP_high_coverage_Illumina.chr${CHR}.filtered.SNV_INDEL_SV_phased_panel.vcf.gz | awk '{gsub(/^chr/,""); print}' | bgzip > ${OUT}.vcf.gz
bcftools index -f ${OUT}.vcf.gz

# BCF
bcftools view ${OUT}.vcf.gz | bcftools view -O b -o ${OUT}.bcf
bcftools index -f ${OUT}.bcf

# BED
plink2 --bcf ${OUT}.bcf --allow-extra-chr --make-bed --out $OUT

# PGEN
plink2 --bcf $OUT.bcf --allow-extra-chr --make-pgen --out $OUT

# BGEN
plink2 --bcf $OUT.bcf --allow-extra-chr --export bgen-1.3 --out $OUT
~/miniconda3/bin/bgenix -g ${OUT}.bgen -index -clobber
```


```{r load, cache=FALSE}
library(GenomicDataStream)
library(ggplot2)
library(tidyverse)
library(peakRAM)

PATH = "~/Downloads/"
```

# Time vs number of variants
```{r numVariants}
file = paste0(PATH, "/1kg_chr21.bed")
k = 20

maf_array = c(0.01, .05, .1, .2, .3, .4)

res = lapply(maf_array, function(MAF){

  gds = GenomicDataStream(file, MAF=MAF)

  res_time = system.time({
    res <- PCAstream(gds, k, threads=4)
  })

  data.frame(MAF = MAF, SNPs = nrow(res$v), t(data.frame(res_time)))
})
res = do.call(rbind, res)
```

```{r numVariantsPlot, cache=FALSE}
ggplot(res, aes(SNPs, elapsed)) +
  geom_point() +
  theme_classic() +
  theme(aspect.ratio=1) +
  ylab("Time (seconds)")
```


# Memory vs number of variants
```{r numVariantsMemory}
maf_array = c(0.01, .05, .1, .2, .3, .4)

res = lapply(maf_array, function(MAF){

  gds = GenomicDataStream(file, MAF=MAF)

  res_mem = peakRAM({
    res <- PCAstream(gds, k, threads=4)
  })

  data.frame(MAF = MAF, SNPs = nrow(res$v), res_mem)
})
res = do.call(rbind, res)
```

```{r numVariantsMemoryPlot, cache=FALSE}
ggplot(res, aes(SNPs, Peak_RAM_Used_MiB / 1e6)) +
  geom_point() +
  theme_classic() +
  theme(aspect.ratio=1) +
  ylab("Memory usage (Gb)")
```


# Time vs number of PCs
```{r numPCs}
file = paste0(PATH, "/1kg_chr21.bed")

k_array = c(2, 5, seq(10, 100, by=10), 
        seq(125, 200, by=25), 
        seq(250, 500, by=50))

res = lapply(k_array, function(k){

  gds = GenomicDataStream(file, MAF=.3)

  res_time = system.time({
    res <- PCAstream(gds, k, threads=4)
  })

  data.frame(k = k, t(data.frame(res_time)))
})
res = do.call(rbind, res)
```

```{r numPCsPlot, cache=FALSE}
ggplot(res, aes(k, elapsed)) +
  geom_point() +
  theme_classic() +
  theme(aspect.ratio=1) +
  ylab("Time (seconds)")
```


# Time vs genotype format
PCA on full file without filtering. Filter is fast with PGEN, BED and BGEN.  But with VCF and BCF, filtering means that each variant is given by its own region.  This is very slow since the variants are zipped by block and requesting two regions within the same block unzips the block each time.



```{r FileFormats, eval=TRUE}
files = dir(PATH, pattern="1kg_chr21", full.names=TRUE)
files = grep(".bcf$|.vcf.gz$|.pgen$|.bed$|.bgen$", files, value=TRUE)

k = 2
res = lapply(files, function(file){

  gds = GenomicDataStream(file, field="GT")

  res_time = system.time({
    res <- PCAstream(gds, k, threads=4)
  })

  data.frame(file, t(data.frame(res_time)))
})
res = do.call(rbind, res)
```


```{r FileFormatsPlot, cache=FALSE, eval=TRUE}
res %>%
  mutate(Type = gsub("1kg_chr21.", "", basename(file))) %>%
  ggplot(aes(Type, elapsed, fill=Type)) +
  geom_bar(stat="identity") +
  theme_classic() +
  theme(aspect.ratio=1, legend.position="none") +
  scale_y_continuous(limits=c(0, NA), expand=c(0,0)) +
  ylab("Time (seconds)") +
  xlab("File type") +
  scale_fill_brewer(palette="Set1")
```



# Time vs number of threads
```{r nThreads}
file = paste0(PATH, "/1kg_chr21.bed")
k = 2
res = lapply(c(1,2,4), function(threads){

  gds = GenomicDataStream(file)

  res_time = system.time({
    res <- PCAstream(gds, k, threads=threads)
  })

  data.frame(threads=threads, t(data.frame(res_time)))
})
res = do.call(rbind, res)
```


```{r nThreadsPlot, cache=FALSE}
ggplot(res, aes(threads, elapsed )) +
  geom_point() +
  theme_classic() +
  theme(aspect.ratio=1, legend.position="none") +
  ylab("Time (seconds)")
```

